Figures from: Vertebrate host availability and the thermal properties of mosquito-borne parasite transmission

### 2023-06-05 manuscript-figures.Rmd

data.R0_norm <- data.R0 %>%
  filter(KH %in% newKH_vals) %>%
  filter(variable == "R0") %>%
  select(-variable) %>%
  # normalize variables to have a maximum of 1
  group_by(system_ID, Model, sigmaH, KH) %>%
  # mutate(mean_norm = mean / max(mean)) %>%
  mutate(median_norm = median / max(median)) %>%
  # mutate(highHCI_norm = highHCI / max(median)) %>%
  # mutate(lowHCI_norm = lowHCI / max(median)) %>%
  arrange(system_ID, sigmaH, KH, Temperature, median_norm)

R0_plot <- ggplot(mapping = aes(x = Temperature, group = KH)) +
  # path of mean, normalized R0 as a function of temperature (finite sigmaH):
  geom_path(
    data = filter(data.R0_norm, sigmaH == 100),
    aes(y = median_norm, colour = KH, linetype = "finite"),
    lwd = 1
  ) +
  # path of mean, normalized R0 as a function of temperature (infinite sigmaH):
  geom_path(
    data = filter(data.R0_norm, sigmaH == Inf),
    aes(y = median_norm, linetype = "infinite", colour = "black"),
    colour = "black",
    lwd = 1.5
  ) +
  # # 89% HCI of R0 TPC curves
  # geom_ribbon(
  #   data = filter(data.R0_norm, sigmaH == 100),
  #   aes(ymin = lowHCI_norm, ymax = highHCI_norm, fill = KH),
  #   alpha = 0.1
  # ) +
  # x-axis:
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0, 0),
    limits = c(14, 34)
  ) +
  # y-axis:
  scale_y_continuous(
    name = TeX("Normalized $R_0$"),
    expand = c(0, .05)
  ) +
  # color: scaled log10, color-blind friendly
  scale_color_viridis(
    name = "Vertebrate host population\ndensity (ind/ha)",
    trans = "log10",
    breaks = 10^seq(-2, 4),
    labels = unname(c(0.01, 0.1, 1, 10, 100, TeX("$10^3$"), TeX("$10^4$"))),
    option = "plasma"
  ) +
  # fill: scaled log10, color-blind friendly
  scale_fill_viridis(
    name = "Vertebrate host population\ndensity (ind/ha)",
    trans = "log10",
    breaks = 10^seq(-2, 4),
    labels = unname(c(0.01, 0.1, 1, 10, 100, TeX("$10^3$"), TeX("$10^4$"))),
    option = "plasma"
  ) +
  # linetype:
  scale_linetype_manual(
    name = "Biting tolerance\n(bites per host per day)",
    breaks = c("finite", "infinite"),
    labels = unname(c(TeX("100 (Chitnis)"), TeX("$\\infty$ (Ross-Macdonald)"))),
    values = c("solid", "22")
  ) +
  # faceting:
  facet_wrap(~system_label, scales = "free", nrow = 2, labeller = labeller()) +
  # legend
  guides(
    color = guide_colourbar(
      title.position = "top",
      nrow = 1,
      barwidth = 12,
      show.limits = TRUE,
      direction = "horizontal"
    ),
    linetype = guide_legend(
      title.position = "top",
      nrow = 2,
      keywidth = 4,
      direction = "vertical"
    )
  ) +
  # theme options:
  theme_minimal_vgrid(16) +
  theme(
    legend.box = "vertical",
    legend.margin = margin(),
    legend.title = element_text(size = 10),
    legend.text = element_text(
      size = 10,
      hjust = 0
    ),
    legend.justification = "left",
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

# Plot
R0_plot <- shift_legend(R0_plot)
```

@@

# add many proxy sigmaH points at infinity to make the values easier to see there
max_sigmaH <- max(filter(Toptalt_df, is.finite(sigmaH))$sigmaH)
# Value of sigmaH to replace 'Inf' with for plotting
infinite_skew <- 10^2.35 # 10^log10(max_sigmaH*1.5) # 1.1*max_sigmaH #10^2 + 80
# Used to create a gap dividing finite and infinite sigmaH
infinite_divide <- 10^2.7 # 10^log10(max_sigmaH*1)

@@

# Plot median of Topt heat map
Toptalt_plots <- Toptalt_df %>%
  ## Set up plot ##
  # x = biting tolerance (with a point at infinity),
  # y = vertebrate host population density,
  # color = transmission thermal optimum
  ggplot(aes(x = KH, y = sigmaH_proxy, z = median)) +
  # Topt for FINITE sigmaH:
  geom_contour_filled(
    data = filter(Toptalt_df, sigmaH_proxy < 10^2.2),
    breaks = Topt_breaks
  ) +
  # Topt for INFINITE sigmaH:
  # take median of median to hide variation from artifacts of uncertainty
  geom_contour_filled(
    data = filter(Toptalt_df, sigmaH_proxy > infinite_skew) %>%
      group_by(system_ID) %>%
      mutate(median = median(median), .groups = "keep"),
    size = 1,
    breaks = Topt_breaks
  ) +
  # # R0 = 1 contour: cheat and use where CTwidth = 0 (i.e. where R0<1 at any temperature)
  geom_contour(
    data = filter(data.CT, variable == "CTwidth", sigmaH < infinite_divide),
    aes(y = sigmaH, z = median, colour = "R0 = 1"),
    breaks = min(filter(data.CT, variable == "CTwidth", median > 0)$median),
    size = 1
  ) +
  # geom_contour(data = data.R0_norm,
  #              aes(y = sigmaH, z = median, colour = "R0=1"), breaks = c(1), size = 1) +
  # x-axis: log10-scale with no buffer space
  scale_x_log10(
    name = TeX("Vertebrate host population density (ind/ha)"),
    expand = c(0, 0),
    limits = c(0.05, 1E4),
    breaks = 10^seq(-1, 4, 1),
    labels = TeX(c("$0.1$", "$1$", "$10", "$100$", "$10^3$", "$10^4$"))
  ) +
  # y-axis: log10-scale y axis, with no buffer space
  scale_y_log10(
    name = TeX("Biting tolerance (bites per host per day)"),
    expand = c(0, 0),
    breaks = sigmaH_breaks,
    labels = sigmaH_labels
  ) +

@@

CTwidth_df <- data.CT %>%
  filter(variable == "CTwidth") %>%
  # filter(sigmaH < 10^2.2) %>%
  mutate(sigmaH_proxy = sigmaH) %>%
  # Arrange along the plotting variables
  arrange(system_ID, KH, sigmaH_proxy, median) %>%
  ungroup()

@@

# Plot
CTWidth_heat_plots <- shift_legend(CTWidth_heat_plots)
```

@@

# !!! [] get enough values of R0 across sigmaH and KH to do a heatmap (21 each?)

@@

# # add a second proxy sigmaH point at infinity to make the values easier to see there
# temp <- filter(R0_df, is.infinite(sigmaH)) %>%
#   mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH, 1.2 * infinite_divide))
# # Create a proxy sigmaH variable that replaces Inf with infinite_skew
# R0_df <- mutate(R0_df, sigmaH_proxy = ifelse(is.finite(sigmaH),
#                                              sigmaH, infinite_skew
# ))
# R0_df <- add_row(R0_df, temp) %>% distinct()

@@

```{r fig-S8_R0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:R0_sens} The sensitivity of the distribution of $\\mathcal{R}_0$ to mosquito life history traits.'}
R0.HPD <- read_rds("results/R0_HPD_sens.rds") %>%
  ungroup() %>%
  pathosystem_labeller()

@@

```{r fig-S9_ddTR0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:ddTR0_sens} The sensitivity of the distribution of the derivative of $\\mathcal{R}_0$ with respect to temperature to mosquito life history traits.'}
ddTR0.HPD <- read_rds("results/ddTR0_HPD_sens.rds") %>%
  ungroup() %>%
  pathosystem_labeller()

@@

plot.ddTR0.rel.sens <- ddTR0.HPD %>%
  filter(KH %in% c(1, 100) & sigmaH %in% c(50, Inf)) %>%
  filter(!(KH == 100 & sigmaH == Inf)) %>%
  filter(rel_HPD_width != 0) %>%
  # For each KH value, sum up all sensitivity values to get total sensitivity across parameters
  group_by(system_ID, sigmaH, KH, Temperature) %>%
  mutate(sens_total = sum(rel_HPD_width)) %>%
  # Divide each sensitivity value by the total sensitivity
  mutate(sens_rel = rel_HPD_width / sens_total) %>%
  arrange(sens_rel) %>%
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = Temperature, y = sens_rel, fill = focal_var)) +
  geom_area() +
  scale_fill_discrete(
    name = "Focal parameter",
    breaks = c("betaV", "deltaL", "etaV", "muV", "rhoL", "sigmaV", "sigmaV_f"),
    labels = c(
      "Vector competence", "Pr(larval survival)",
      "Parasite development rate", "Mortality rate",
      "Immature development rate", "Max. biting rate",
      "Eggs per female per day"
    )
  ) +
  scale_x_continuous(
    limits = Temp_range_for_plot,
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    name = unname(TeX("Relative contribution to $dR_0/dT$ HPD width"))
  ) +
  facet_grid(
    cols = vars(system_label),
    rows = vars(KH, sigmaH),
    labeller = labeller(
      system_label = label_wrap_gen(width = 10),
      sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
      KH = as_labeller(appender_KH, default = label_parsed)
    ),
    scales = "free_y"
  ) +
  theme_minimal_hgrid(16) +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
plot.ddTR0.rel.sens
```
```{r fig-S10_Topt_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_sens} The sensitivity of the thermal optimmum to mosquito life history traits as a function of vertebrate host population density when $\\sigma_H = 50$.'}
Topt.HPD <- read_rds("results/Topt_HPD_sens.rds") %>%
  ungroup() %>%
  pathosystem_labeller() %>% 
  filter(!is.na(rel_HPD_width)) %>%
  # For each KH value, sum up all sensitivity values to get total sensitivity
  group_by(system_ID, sigmaH, KH) %>%
  mutate(sens_total = sum(rel_HPD_width)) %>%
  # Divide each sensitivity value by the total sensitivity
  mutate(sens_rel = rel_HPD_width / sens_total) # %>%
# arrange(sens_rel)

>>>>

data.R0_norm <- data.R0 %>%
  filter(KH %in% newKH_vals) %>%
  filter(variable == "R0") %>%
  select(-variable) %>%
  # normalize median of R0 distribution to have a maximum of one
  group_by(system_ID, Model, sigmaH, KH) %>%
  mutate(median_norm = median / max(median)) %>%
  arrange(system_ID, sigmaH, KH, Temperature, median_norm)

R0_plot <- ggplot(mapping = aes(x = Temperature, group = KH)) +
  # path of mean, normalized R0 as a function of temperature (finite sigmaH):
  geom_path(
    data = filter(data.R0_norm, sigmaH == 100),
    aes(y = median_norm, colour = KH, linetype = "finite"),
    lwd = 1
  ) +
  # path of mean, normalized R0 as a function of temperature (infinite sigmaH):
  geom_path(
    data = filter(data.R0_norm, sigmaH == Inf),
    aes(y = median_norm, linetype = "infinite", colour = "black"),
    colour = "black",
    lwd = 1.5
  ) +
  # x-axis:
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0, 0),
    limits = c(14, 34)
  ) +
  # y-axis:
  scale_y_continuous(
    name = TeX("Normalized $R_0$"),
    expand = c(0, .05)
  ) +
  # color: scaled log10, color-blind friendly
  scale_color_viridis(
    name = "Vertebrate host population\ndensity (ind/ha)",
    trans = "log10",
    breaks = 10^seq(-2, 4),
    labels = unname(c(0.01, 0.1, 1, 10, 100, TeX("$10^3$"), TeX("$10^4$"))),
    option = "plasma"
  ) +
  # fill: scaled log10, color-blind friendly
  scale_fill_viridis(
    name = "Vertebrate host population\ndensity (ind/ha)",
    trans = "log10",
    breaks = 10^seq(-2, 4),
    labels = unname(c(0.01, 0.1, 1, 10, 100, TeX("$10^3$"), TeX("$10^4$"))),
    option = "plasma"
  ) +
  # linetype:
  scale_linetype_manual(
    name = "Biting tolerance\n(bites per host per day)",
    breaks = c("finite", "infinite"),
    labels = unname(c(TeX("100 (Chitnis)"), TeX("$\\infty$ (Ross-Macdonald)"))),
    values = c("solid", "22")
  ) +
  # faceting:
  facet_wrap(~system_label, scales = "free", nrow = 2, labeller = labeller()) +
  # legend
  guides(
    color = guide_colourbar(
      title.position = "top",
      nrow = 1,
      barwidth = 12,
      show.limits = TRUE,
      direction = "horizontal"
    ),
    linetype = guide_legend(
      title.position = "top",
      nrow = 2,
      keywidth = 4,
      direction = "vertical"
    )
  ) +
  # theme options:
  theme_minimal_vgrid(16) +
  theme(
    legend.box = "vertical",
    legend.margin = margin(),
    legend.title = element_text(size = 10),
    legend.text = element_text(
      size = 10,
      hjust = 0
    ),
    legend.justification = "left",
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

# Plot
shift_legend(R0_plot)
```

@@

# add many proxy sigmaH points at infinity to make the values easier to see there
max_sigmaH <- max(filter(Toptalt_df, is.finite(sigmaH))$sigmaH)
# Value of sigmaH to replace 'Inf' with for plotting
infinite_skew <- 10^2.35
# Used to create a gap dividing finite and infinite sigmaH
infinite_divide <- 10^2.7 # 10^log10(max_sigmaH*1)

@@

# Plot median of Topt heat map
Toptalt_plots <- Toptalt_df %>%
  ## Set up plot ##
  # x = biting tolerance (with a point at infinity),
  # y = vertebrate host population density,
  # color = transmission thermal optimum
  ggplot(aes(x = KH, y = sigmaH_proxy, z = median)) +
  # Topt for FINITE sigmaH:
  geom_contour_filled(
    data = filter(Toptalt_df, sigmaH_proxy < 10^2.2),
    breaks = Topt_breaks
  ) +
  # Topt for INFINITE sigmaH:
  # take median of median to hide variation from artifacts of uncertainty
  geom_contour_filled(
    data = filter(Toptalt_df, sigmaH_proxy > infinite_skew) %>%
      group_by(system_ID) %>%
      mutate(median = median(median), .groups = "keep"),
    size = 1,
    breaks = Topt_breaks
  ) +
  # # R0 = 1 contour: cheat and use where CTwidth = 0 (i.e. where R0<1 at any temperature)
  geom_contour(
    data = filter(data.CT, variable == "CTwidth", sigmaH < infinite_divide),
    aes(y = sigmaH, z = median, colour = "R0 = 1"),
    breaks = min(filter(data.CT, variable == "CTwidth", median > 0)$median),
    size = 1
  ) +
  # x-axis: log10-scale with no buffer space
  scale_x_log10(
    name = TeX("Vertebrate host population density (ind/ha)"),
    expand = c(0, 0),
    limits = c(0.05, 1E4),
    breaks = 10^seq(-1, 4, 1),
    labels = TeX(c("$0.1$", "$1$", "$10", "$100$", "$10^3$", "$10^4$"))
  ) +
  # y-axis: log10-scale y axis, with no buffer space
  scale_y_log10(
    name = TeX("Biting tolerance (bites per host per day)"),
    expand = c(0, 0),
    breaks = sigmaH_breaks,
    labels = sigmaH_labels
  ) +

@@

CTwidth_df <- data.CT %>%
  filter(variable == "CTwidth") %>%
  mutate(sigmaH_proxy = sigmaH) %>%
  # Arrange along the plotting variables
  arrange(system_ID, KH, sigmaH_proxy, median) %>%
  ungroup()

@@

# Plot
shift_legend(CTWidth_heat_plots)
```

@@

@@

@@

```{r fig-SX_R0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:R0_sens} The sensitivity of the distribution of $\\mathcal{R}_0$ to mosquito life history traits.'}

# !!! [] be sure to use a consistent list of mosquito life history traits:
#         - adult lifespan (1/mu_V)
#         - biting frequency (sigma_V)
#         - eggs per female per day (sigma_V * f)
#         - immature development rate
#         - Pr(larval survival) (deltaL)
#         - Pr(surviving EIP) (theta_V)
#         - vector competence (beta_V)

R0.HPD <- read_rds("results/R0_HPD_sens.rds") %>%
  ungroup() %>%
  pathosystem_labeller()

@@

```{r fig-SX_ddTR0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:ddTR0_sens} The sensitivity of the distribution of the derivative of $\\mathcal{R}_0$ with respect to temperature to mosquito life history traits.'}

# !!! [] be sure to use a consistent list of mosquito life history traits:
#         - adult lifespan (1/mu_V)
#         - biting frequency (sigma_V)
#         - eggs per female per day (sigma_V * f)
#         - immature development rate
#         - Pr(larval survival) (deltaL)
#         - Pr(surviving EIP) (theta_V)
#         - vector competence (beta_V)

ddTR0.HPD <- read_rds("results/ddTR0_HPD_sens.rds") %>%
  ungroup() %>%
  pathosystem_labeller()

@@

plot.ddTR0.rel.sens <- ddTR0.HPD %>%
  filter(KH %in% c(1, 100) & sigmaH %in% c(50, Inf)) %>%
  filter(!(KH == 100 & sigmaH == Inf)) %>%
  filter(rel_HPD_width != 0) %>%
  # For each KH value, sum up all sensitivity values to get total sensitivity across parameters
  group_by(system_ID, sigmaH, KH, Temperature) %>%
  mutate(sens_total = sum(rel_HPD_width)) %>%
  # Divide each sensitivity value by the total sensitivity
  mutate(sens_rel = rel_HPD_width / sens_total) %>%
  arrange(sens_rel) %>%
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = Temperature, y = sens_rel, fill = focal_var)) +
  geom_area() +
  scale_fill_discrete(
    name = "Focal parameter",
    breaks = c("betaV", "deltaL", "etaV", "muV", "rhoL", "sigmaV", "sigmaV_f"),
    labels = c(
      "Vector competence", "Pr(larval survival)",
      "Parasite development rate", "Mortality rate",
      "Immature development rate", "Max. biting rate",
      "Eggs per female per day"
    )
  ) +
  scale_x_continuous(
    limits = Temp_range_for_plot,
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    name = unname(TeX("Relative contribution to $dR_0/dT$ HPD width"))
  ) +
  facet_grid(
    cols = vars(system_label),
    rows = vars(KH, sigmaH),
    labeller = labeller(
      system_label = label_wrap_gen(width = 10),
      sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
      KH = as_labeller(appender_KH, default = label_parsed)
    ),
    scales = "free_y"
  ) +
  theme_minimal_hgrid(16) +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
plot.ddTR0.rel.sens
```

```{r fig-SX_Topt_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_sens} The sensitivity of the thermal optimmum to mosquito life history traits as a function of vertebrate host population density when $\\sigma_H = 50$.'}

# !!! [] be sure to use a consistent list of mosquito life history traits:
#         - adult lifespan (1/mu_V)
#         - biting frequency (sigma_V)
#         - eggs per female per day (sigma_V * f)
#         - immature development rate
#         - Pr(larval survival) (deltaL)
#         - Pr(surviving EIP) (theta_V)
#         - vector competence (beta_V)

Topt.HPD <- read_rds("results/Topt_HPD_sens.rds") %>%
  ungroup() %>%
  pathosystem_labeller() %>%
  filter(!is.na(rel_HPD_width)) %>%
  # For each KH value, sum up all sensitivity values to get total sensitivity
  group_by(system_ID, sigmaH, KH) %>%
  mutate(sens_total = sum(rel_HPD_width)) %>%
  # Divide each sensitivity value by the total sensitivity
  mutate(sens_rel = rel_HPD_width / sens_total) # %>%
# arrange(sens_rel)
